---
title: "Build Data Set"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook builds the analytical data set from the results of `redlining_area.Rmd`, CDC data, and demographic data.

## Dependencies
This notebook requires a number of packages:

```{r load-packages}
# tidyverse packages
library(dplyr)         # data wrangling

# other packages
library(cityHealth)    # CDC data
library(here)          # file path management
library(tidycensus)    # demographic data
```

## Load Data
This notebook requires the redlining data already calculated for St. Louis:

```{r load-data}
redlining <- read_csv(here("data", "redlining", "STL_REDLINING_cdGrade.csv"))
```

## Demographic Data

The first goal of this notebook is to download and prepare for joining all of the needed demographic data.

### Race
*individuals*
```{r}
race <- get_acs(geography = "tract", table = "B02001", state = 29, county = 510, 
                     year = 2016, output = "wide")

race %>%
  rename(totalPop = B02001_001E) %>%
  rename(totalPop_m = B02001_001M) %>%
  rename(white = B02001_002E) %>%
  rename(white_m = B02001_002M) %>%
  rename(black = B02001_003E) %>%
  rename(black_m = B02001_003M) %>%
  mutate(otherRace = totalPop-white-black) %>%
  select(GEOID, totalPop, totalPop_m, white, white_m, black, black_m, otherRace) -> race
```

### Median Income
*household*
```{r}
medianInc <- get_acs(geography = "tract", variables = "B19019_001", state = 29, county = 510, 
                     year = 2016, output = "wide")

medianInc %>%
  select(-NAME) %>%
  rename(medianInc = B19019_001E) %>%
  rename(medianInc_m = B19019_001M) -> medianInc
```

### Poverty Status
*individuals*
```{r}
poverty <- get_acs(geography = "tract", variables = "B17001_002", state = 29, county = 510, 
                     year = 2016, output = "wide")

poverty %>%
  select(-NAME) %>%
  rename(poverty = B17001_002E) %>%
  rename(poverty_m = B17001_002M) -> poverty
```

*individuals*
```{r}
povertyB <- get_acs(geography = "tract", variables = "B17001B_002", state = 29, county = 510, 
                     year = 2016, output = "wide")

povertyB %>%
  select(-NAME) %>%
  rename(povertyB = B17001B_002E) %>%
  rename(povertyB_m = B17001B_002M) -> povertyB
```

*individuals*


```{r}
medicaid <- get_acs(geography = "tract", table = "C27007", state = 29, county = 510, 
                     year = 2016, output = "wide")

medicaid %>%
  mutate(medicaid = C27007_004E + C27007_007E + C27007_010E + C27007_014E + C27007_017E + C27007_020E) %>%
  select(GEOID, medicaid) -> medicaid
```


*households*


```{r}
tanf <- get_acs(geography = "tract", table = "B19057", state = 29, county = 510, 
                     year = 2016, output = "wide")

tanf %>%
  rename(house = B19057_001E) %>%
  rename(house_m = B19057_001M) %>%
  rename(tanf = B19057_002E) %>%
  rename(tanf_m = B19057_002M) %>%
  mutate(tanfProp = tanf/house) %>%
  select(GEOID, house, house_m, tanf, tanf_m, tanfProp) -> tanf
```

*households* 
```{r}
snap <- get_acs(geography = "tract", table = "B19058", state = 29, county = 510, 
                     year = 2016, output = "wide")

snap %>%
  rename(snap = B19058_002E) %>%
  rename(snap_m = B19058_002M) %>%
  mutate(snapProp = snap/B19058_001E) %>%
  select(GEOID, snap, snap_m, snapProp) -> snap
```

*households*
 
```{r}
tenure <- get_acs(geography = "tract", table = "B25003", state = 29, county = 510, 
                     year = 2016, output = "wide")

tenure %>%
  rename(own = B25003_002E) %>%
  rename(own_m = B25003_002M) %>%
  mutate(ownProp = own/B25003_001E) %>%
  rename(rent = B25003_003E) %>%
  rename(rent_m = B25003_003M) %>%
  mutate(rentProp = rent/B25003_001E) %>%
  select(GEOID, own, own_m, ownProp, rent, rent_m, rentProp) -> tenure
```

## CDC Data

uninsurance rate




