---
title: "Map and Plot Data"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook creates the analytical maps and plots used in my presentation of the pilot work in Boston in October 2018.

## Dependencies
This notebook requires a number of packages for plotting data:

```{r load-packages}
# tidyverse packages
library(dplyr)       # data wrangling
library(ggplot2)     # data plotting

# other packages
library(here)        # file path management
library(prener)      # data plotting
library(sf)          # mapping
```

## Load Data
This notebook requires the analysis shapefile created by `build_data.Rmd`:

```{r load-data}
tracts <- st_read(here("data", "spatial", "clean", "STL_REDLINING_Analysis.shp"), stringsAsFactors = FALSE)

load(here("data", "spatial", "baseMap", "baseMap.rda"))
```

## Update CRS for Tract Data
We want both data sets to have matching `crs` values - the tract data are therefore updates to EPSG `4326`:

```{r update-crs}
city <- st_transform(city, crs = 4326)
city_hwys <- st_transform(city_hwys, crs = 4326)
tracts <- st_transform(tracts, crs = 4326)
```

## Additional Data Cleaning

```{r}
tracts <- mutate(tracts, CDPROP = CDGRADE/as.numeric(ALAND))
```


## Race and Redlining

```{r}
dissimJenks <- cp_breaks(tracts, var = DISSIM, newvar = DJENKS, classes = 5, style = "jenks")

ggplot() +
  geom_sf(data = dissimJenks, mapping = aes(fill = DJENKS), color = "#A6AAA9") +
  geom_sf(data = city_hwys, mapping = aes(color = "Highways"), size = 1.5, fill = NA) +
  geom_sf(data = city, fill = NA, color = "#000000", size = .25) +
  scale_fill_brewer(palette = "RdBu", name = "Dissimilarity",
                    labels = c("< -0.016", "-0.016 - -0.004", "-0.004 - 0.009", "0.009 - 0.021", "> 0.021")) +
  scale_colour_manual(name="", values= "black") +
  labs(
    title = " ",
    subtitle = "Segregation"
  ) +
  cp_sequoiaTheme(background = "transparent", map = TRUE)

# save plot
cp_plotSave(filename = here("results", "maps", "dissim.png"), preset = "lg", dpi = 500)

```


```{r}
subtitle <- expression(paste("Pearson's ", italic("r"), " = 0.286, ", italic("p"), " = .003"))

ggplot(data = tracts, mapping = aes(x = BLACK, y = CDPROP)) + 
  geom_point(position = "jitter", shape = 21, fill = "#349E8B", size = 4) + 
  geom_smooth(method = lm, color = "#E16033", size = 2) +
  labs(
    title = "Redlining and Race per Census Tract",
    subtitle = subtitle,
    x = "Proportion African American",
    y = "Proportion C or D Grade"
  ) +
  cp_sequoiaTheme(background = "gray")

# save plot
cp_plotSave(filename = here("results", "figures", "redliningRace.png"), preset = "lg", dpi = 500)
```


```{r}
subtitle <- expression(paste("Pearson's ", italic("r"), " = -0.272, ", italic("p"), " = 0.005"))

ggplot(data = tracts, mapping = aes(x = DISSIM, y = CDPROP)) + 
  geom_point(position = "jitter", shape = 21, fill = "#349E8B", size = 4) + 
  geom_smooth(method = lm, color = "#E16033", size = 2) +
  labs(
    title = "Redlining and Dissimilarity per Census Tract",
    subtitle = subtitle,
    x = "Index of Dissimilarity",
    y = "Proportion C or D Grade",
    caption = "Negative dissimilarity values indicate segregated black neighborhoods."
  ) +
  cp_sequoiaTheme(background = "gray")

# save plot
cp_plotSave(filename = here("results", "figures", "redliningDissimilarity.png"), preset = "lg", dpi = 500)
```



```{r}
subtitle <- expression(paste("Pearson's ", italic("r"), " = 0.432, ", italic("p"), " < 0.001"))

ggplot(data = tracts, mapping = aes(x = POVERTY, y = CDPROP)) + 
  geom_point(position = "jitter", shape = 21, fill = "#349E8B", size = 4) + 
  geom_smooth(method = lm, color = "#E16033", size = 2) +
  labs(
    title = "Redlining and Poverty per Census Tract",
    subtitle = subtitle,
    x = "Proportion of Individuals Below Poverty Line",
    y = "Proportion C or D Grade"
  ) +
  cp_sequoiaTheme(background = "gray")

# save plot
cp_plotSave(filename = here("results", "figures", "redliningPoverty.png"), preset = "lg", dpi = 500)
```


```{r}
subtitle <- expression(paste("Pearson's ", italic("r"), " = 0.328, ", italic("p"), " = 0.001"))

ggplot(data = tracts, mapping = aes(x = ASTHMA, y = CDPROP)) + 
  geom_point(position = "jitter", shape = 21, fill = "#349E8B", size = 4) + 
  geom_smooth(method = lm, color = "#E16033", size = 2) +
  labs(
    title = "Redlining and Asthma per Census Tract",
    subtitle = subtitle,
    x = "Proportion of Individuals with Current Asthma",
    y = "Proportion C or D Grade"
  ) +
  cp_sequoiaTheme(background = "gray")

# save plot
cp_plotSave(filename = here("results", "figures", "redliningAsthma.png"), preset = "lg", dpi = 500)
```


```{r}
asthmaJenks <- cp_breaks(tracts, var = ASTHMA, newvar = AJENKS, classes = 5, style = "jenks")

ggplot() +
  geom_sf(data = asthmaJenks, mapping = aes(fill = AJENKS), color = "#A6AAA9") +
  geom_sf(data = city_hwys, mapping = aes(color = "Highways"), size = 1.5, fill = NA) +
  geom_sf(data = city, fill = NA, color = "#000000", size = .25) +
  scale_fill_brewer(palette = "Blues", name = "Proportion",
                    labels = c("< 0.094", "0.083 - 0.106", "0.106 - 0.121", "0.121 - 0.134", "0.134 - 0.153")) +
  scale_colour_manual(name="", values= "black") +
  labs(
    title = "Current Asthma Diagnosis"
  ) +
  cp_sequoiaTheme(background = "transparent", map = TRUE)

# save plot
cp_plotSave(filename = here("results", "maps", "asthma.png"), preset = "lg", dpi = 500)
```


```{r}
subtitle <- expression(paste("Pearson's ", italic("r"), " = 0.900, ", italic("p"), " < 0.001"))

ggplot(data = tracts, mapping = aes(x = ASTHMA, y = BLACK)) + 
  geom_point(position = "jitter", shape = 21, fill = "#349E8B", size = 4) + 
  geom_smooth(method = lm, color = "#E16033", size = 2) +
  labs(
    title = "Race and Asthma per Census Tract",
    subtitle = subtitle,
    x = "Proportion of Individuals with Current Asthma",
    y = "Proportion African Americans"
  ) +
  cp_sequoiaTheme(background = "gray")

# save plot
cp_plotSave(filename = here("results", "figures", "raceAsthma.png"), preset = "lg", dpi = 500)
```
