---
title: "Calculate Redlining Area"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook creates measures of the impact of redlining by census tract. Two measures are calculated - the square meters of Class 'C' (yellow, or declining areas) and the square meters of Class 'D' (red, or minority/immigrant areas).

## Dependencies
This notebook requires data from the `data/` directory as well as the `sf` and `dplyr` packages.

```{r load-packages}
# project package
library(redHealth)

# tidyverse packages
library(dplyr)     # data wrangling

# spatial packages
library(sf)        # spatial data tools

# other packages
library(here)      # file path management
```

## Load Data
This notebook requires the raw redlining data and the census tract boundary data.

```{r load-data}
redRaw <- st_read(here("data", "redlining", "raw", "HOLC_St.shp"), stringsAsFactors = FALSE)
tracts <- st_read(here("data", "spatial", "raw", "STL_DEMOGRAPHICS_Tracts.shp"), stringsAsFactors = FALSE)
```

## Update CRS for Tract Data
We want both data sets to have matching `crs` values - the tract data are therefore updates to EPSG `4326`:

```{r update-crs}
tracts <- st_transform(tracts, crs = 4326)
```

They can now be used for geometric operations with the `redRaw` data.

## Clean Redlining Data
The first step is to remove unneeded columns and unneeded observations from the `redRaw` sf object:

```{r subset-columns}
redRaw %>% 
  select(-city, -name) %>%
  filter(holc_grade == "C" | holc_grade == "D") -> cdAreas
```

Next, we want to calculate the area of each census tract that is covered by both "C" and "D" graded zones:

```{r calculate-area}
cAreas <- rh_area(tract = tracts, holc = cdAreas, cat = "C")
dAreas <- rh_area(tract = tracts, holc = cdAreas, cat = "D")
```

```{r}
redAreas <- rh_area_join(tract = tracts, area = cAreas, by = "GEOID", cat = "cGrade")
redAreas <- rh_area_join(tract = redAreas, area = dAreas, by = "GEOID", cat = "dGrade")
```

